---
title: "Potato Soil Health"
author: "Sudha GC Upadhaya"
date: "2023-01-21"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


```{r, echo=T, warning = F, cache=FALSE, }
library(ggplot2)
library(raster)
library(agricolae)
library(tidyr)
library(ggpubr)
library(dplyr)
library(tidyverse)
library(data.table)
library(vegan)
library(lme4)
library(lmerTest)
```

# Plot sampling locations on Map

```{r, warning = F}
df = read_csv(file = 'C:/Users/theox/Desktop/Projects/SoilFingerPrints/Data/SamplingLocations/SamplingLocations.csv')
df = df[c('ID', 'lat', 'long', 'FieldType')]
head(df)
df$FieldType = gsub('Field', "", df$FieldType)
df$FieldType = gsub('field', "", df$FieldType)
df$FieldType = gsub('Non-virgin ', "Non-virgin", df$FieldType)
df$FieldType = gsub('Virgin ', "Virgin", df$FieldType)
unique(df$FieldType)
```


```{r, warning = F}
state_list    <- c('California', 'Nevada', 'Utah', 'Wyoming', 
                   'Montana', 'Idaho', 'Oregon', 'Washington')
province_list <- c("British Columbia", "Alberta")
us <- raster::getData("GADM",country="USA",level=1)
canada <- raster::getData("GADM",country="CAN",level=1)
us.states <- us[us$NAME_1 %in% state_list,]
ca.provinces <- canada[canada$NAME_1 %in% province_list,]
```


```{r, warning = F}
options(repr.plot.width=10, repr.plot.height=8)
#jpeg('Map_soilprints.jpg',width=10,height=8,units='in',res=300)
Map = ggplot(us.states,aes(x=long,y=lat, group = group))+
  geom_path(size = 0.25)+
  geom_path(data=ca.provinces, size = 0.25, color = 'grey60')+
  geom_path(data=subset(us.states, NAME_1 %in% c('Montana', 'California','Nevada','Utah','Wyoming')),
            size=0.25,color="grey60")+
  geom_polygon(data=subset(us.states, NAME_1 %in% c('Idaho', 'Oregon', 'Washington')),
               aes(x=long, y=lat, group=group), fill="white", colour = "black", size=0.50)+
  theme(panel.background = element_rect(fill = "white", colour = "grey50"))+
  coord_map("bonne", lat0 = 50,
            xlim = c(-124.5, -116.5),ylim = c(44, 49))+
  xlab("Longitude") +
  ylab("Latitude")+
  geom_point(data = df, aes(x= long, y = lat,color = FieldType, fill = FieldType), position=position_jitter(h=0.07, w=0.07),shape =21, size =6, group = FALSE, alpha =0.5)+
  scale_fill_manual(name = 'Field Type', labels = c('Non-virgin', 'Virgin', 'Native'), values = c('orange', 'chartreuse4', 'tomato'))+
  scale_colour_manual(name = 'Field Type', labels = c('Non-virgin', 'Virgin', 'Native'),values = c('black', 'black', 'black')) +
  guides(color = guide_legend('Field Type'))+
  theme_classic()+
  theme(axis.title.x= element_text(size =18, color ='black'),
        axis.title.y = element_text(size =18, color ='black'),
        axis.text.x = element_text(size =15,color = 'black'),
        axis.text.y = element_text(size =15, color = 'black'),
        legend.text = element_text(size =15),
        legend.title = element_text(size =17),
  )
Map

#dev.off()
```

# 2021 data visulatization and inference

```{r}
df_21 = read.csv('C:/Users/theox/Desktop/Projects/SoilFingerPrints/Data/2021/SoilFingerPrintsData_21.csv')
df_21[c("ID", "Pair", "FieldType","Block")] <- lapply(df_21[c("ID", "Pair", "FieldType", "Block")],
                                                factor)
df_21 = df_21[, c(1, 3, 4:18, 20, 22, 23,24,26:32, 34)]
#lapply(df, unique)
head(df_21)
```
## Microplot study

### Senescence

#### AUSPC calcuation

```{r}
library('agricolae')
ts = gsub("(?:[^.]+\\.){1}([^.]+).*", "\\1",
    names(df_21)[(grepl("Sen", names(df_21)))])
ts = as.numeric(ts)
ts[2:5] = ts[2:5] +31
ts[6:8] = ts[6:8]+60#= ts[2]+30
ts = ts-28
ts
names(df_21)[(grepl("Sen", names(df_21)))] <- ts
df_21$AUDPC <- audpc(df_21[,9:16],ts)
head(df_21$AUDPC)
names(df_21)
```

#### Plot Senescence

```{r}
options(repr.plot.width = 5, repr.plot.height = 3)
#jpeg('Sen_21.jpg',width=5,height=3,units='in',res=300)
c1= ggplot(data = df_21, aes(x =FieldType, y = AUDPC, color = FieldType))+
    geom_boxplot(width = 0.4)+
    geom_jitter(alpha=0.8, pch =20,size =5, position=position_jitter(width=0.1))+
    theme_bw()+
    labs(x = 'Field Type', y="Senescence (AUSPC)") +
    theme(legend.position = "none")+
    theme(axis.title.x= element_blank(),
        axis.title.y = element_text(size =10, color ='black'),
        axis.text.x = element_text(size =10, color ='black'),
        axis.text.y = element_text(size =8, color = 'black'))+
scale_color_manual(name = 'Field Type', labels = c('Non-virgin', 'Virgin'), values = c('orange', 'chartreuse4'))+
  facet_wrap(~ Location)
#c1
#dev.off()
```


### Tuber weight (more than 20 gm)

```{r}
options(repr.plot.width = 5, repr.plot.height = 3)
#jpeg('TuberwtAll_soilprints.jpg',width=5,height=3,units='in',res=300)
a1= ggplot(data = df_21, aes(x =FieldType, y = Weight_GT20, color = FieldType))+
    geom_boxplot(width = 0.4)+
    geom_jitter(alpha=0.8, pch =20,size =5, position=position_jitter(width=0.1))+
    theme_bw()+
    labs(x = 'Field Type', y="Tuber weight (gm)") +
    theme(legend.position = "none")+
    theme(axis.title.x= element_blank(),
        axis.title.y = element_text(size =10, color ='black'),
        axis.text.x = element_text(size =10, color ='black'),
        axis.text.y = element_text(size =8, color = 'black'))+
scale_color_manual(name = 'Field Type', labels = c('Non-virgin', 'Virgin'), values = c('orange', 'chartreuse4'))+
  facet_wrap(~ Location)
#a1
#dev.off()
```

### Tuber count (more than 20 gm)

```{r}
options(repr.plot.width = 5, repr.plot.height = 3)
#jpeg('TuberwtAll_soilprints.jpg',width=5,height=3,units='in',res=300)
b1= ggplot(data = df_21, aes(x =FieldType, y = Count_GT20, color = FieldType))+
    geom_boxplot(width = 0.4)+
    geom_jitter(alpha=0.8, pch =20,size =5, position=position_jitter(width=0.1))+
    theme_bw()+
    labs(x = 'Field Type', y="Number of tubers") +
    theme(legend.position = "none")+
    theme(axis.title.x= element_blank(),
        axis.title.y = element_text(size =10, color ='black'),
        axis.text.x = element_text(size =10, color ='black'),
        axis.text.y = element_text(size =8, color = 'black'))+
scale_color_manual(name = 'Field Type', labels = c('Non-virgin', 'Virgin'), values = c('orange', 'chartreuse4'))+
  facet_wrap(~ Location)
#b1
#dev.off()
```


```{r}
options(repr.plot.width = 10, repr.plot.height = 5)
#jpeg('microplot_21.jpg',width=10,height=5,units='in',res=300)
microplot_21 = ggarrange(a1,b1,c1, labels = c("A", "B", "C"), ncol = 2, nrow = 2)
microplot_21
#dev.off()
```



#### Calculate yield Difference 

```{r}
df21_wt_diff = df_21 %>% dplyr::select(FieldType, Pair, Wt_Tuber.g., Location) %>% 
  spread(FieldType, Wt_Tuber.g.)
df21_wt_diff = rename(df21_wt_diff, non_virgin = `Non-virgin field`, virgin = `Virgin field`)
df21_wt_diff = mutate(df21_wt_diff, diff = virgin - non_virgin)
df21_wt_diff
```

#### plot yield difference

```{r, echo = F, include = F}
options(repr.plot.width = 2, repr.plot.height = 2)
#jpeg('yield_21_difference.jpg',width=3,height=3,units='in',res=300)
yield_difference = ggplot(data = df21_wt_diff, aes(x =Location, y = diff, color = Location))+
    geom_boxplot(width = 0.4)+geom_hline(yintercept = 0)+ ylim(-1500, 1500)+
    geom_jitter(alpha=0.8, pch =20,size =5, position=position_jitter(width=0.1, height = 0.3))+
    theme_bw()+
    labs(x = 'Location', y=expression(paste("Difference (Tuber yield)"))) +
    theme(legend.position = "none")+
    theme(axis.title.x= element_blank(),
        axis.title.y = element_text(size =10, color ='black'),
        axis.text.x = element_text(size =10, color ='black'),
        axis.text.y = element_text(size =8, color = 'black'))+
scale_color_manual(name = 'Location', labels = c('Columbia Basin', 'Skagit Valley'), 
                  values = c('darkgoldenrod1', 'darkgreen'))
yield_difference
#dev.off()
```

#### Inferece on weight

```{r}
fit_yield21 = lmer(Wt_Tuber.g. ~ FieldType + (1|Location), 
                  data = df_21)
summary(fit_yield21)
anova(fit_yield21)

## for columbia basin
df_21_CB = df_21[df_21$Location == 'Columbia Basin', ]
anova(lmer(Wt_Tuber.g. ~ FieldType + (1|Pair), 
                  data = df_21_CB))
anova(lm(Wt_Tuber.g. ~ FieldType, data = df_21_CB))

# conduct paired t test
t.test(Wt_Tuber.g.~ FieldType, 
       #paired=T, 
       data=df_21_CB)

#for skagit valley
df_21_SV = df_21[df_21$Location == 'Skagit Valley', ]
anova(lmer(Wt_Tuber.g. ~ FieldType + (1|Pair), 
                  data = df_21_SV))
anova(lm(Wt_Tuber.g. ~ FieldType, data = df_21_SV))

# conduct paired t test
t.test(Wt_Tuber.g.~ FieldType, 
       paired=T, 
       data=df_21_SV)


# Test if there is difference between locations
t.test(diff~ Location, 
       data=df21_wt_diff)

```

#### Inference on all variables

```{r}
inf = function(x ,df = df_21){
  fit = lmer(x ~ FieldType + (1|Location), 
                  data = df)
return(anova(fit))}

variables = names(df_21)[-c(1,2,3,4,5,10,11,12,13,14,15,16,17,18,19, 20,21,25)]

df_var = df_21[variables]

for (i in 1:length(df_var)){
  print(names(df_var)[i])
  f = inf(df_var[[i]], df_21)
  print(f)
}

```

#### Columbia Basin

```{r}
df_21_CB = df_21[df_21$Location == 'Columbia Basin', ]
inf = function(x ,df = df_21_CB){
  fit = lmer(x ~ FieldType + (1|Pair), 
                  data = df)
return(anova(fit))}

variables = names(df_21_CB)[-c(1,2,3,4,5,10,11,12,13,14,15,16,17,18,19, 20,21,25)]
variables
df_var = df_21_CB[variables]

for (i in 1:length(df_var)){
  print(names(df_var)[i])
  f = inf(df_var[[i]], df_21_CB)
  print(f)
}

```


#### Skagit valley

```{r, eval = F}
df_21_SV = df_21[df_21$Location == 'Skagit Valley', ]
inf = function(x ,df = df_21_SV){
  fit = lmer(x ~ FieldType + (1|Pair), 
                  data = df)
return(anova(fit))}

variables = names(df_21_SV)[-c(1,2,3,4,5,10,11,12,13,14,15,16,17,18,19, 20,21,25, 28)]
variables
df_var = df_21_SV[variables]

for (i in 1:length(df_var)){
  print(names(df_var)[i])
  f = inf(df_var[[i]], df_21_SV)
  print(f)
}

```


## Soil biological properties 

### Pythium 

```{r}
options(repr.plot.width = 5, repr.plot.height = 3)
#jpeg('TuberwtAll_soilprints.jpg',width=5,height=3,units='in',res=300)
sb1= ggplot(data = df_21, aes(x =FieldType, y = Pyth_Avg_Soil, color = FieldType))+
    geom_boxplot(width = 0.4)+
    geom_jitter(alpha=0.8, pch =20,size =5, position=position_jitter(width=0.1))+
    theme_bw()+
    labs(x = 'Field Type', y=expression(paste("Average ",italic("Pythium")))) +
    theme(legend.position = "none")+
    theme(axis.title.x= element_blank(),
        axis.title.y = element_text(size =10, color ='black'),
        axis.text.x = element_text(size =10, color ='black'),
        axis.text.y = element_text(size =8, color = 'black'))+
scale_color_manual(name = 'Field Type', labels = c('Non-virgin', 'Virgin'), values = c('orange', 'chartreuse4'))+
  facet_wrap(~ Location)
#sb1
#dev.off()
```

### Resistant Pythium

```{r}
options(repr.plot.width = 5, repr.plot.height = 3)
#jpeg('TuberwtAll_soilprints.jpg',width=5,height=3,units='in',res=300)
sb2= ggplot(data = df_21, aes(x =FieldType, y = ResPyth_Avg_Soil, color = FieldType))+
    geom_boxplot(width = 0.4)+
    geom_jitter(alpha=0.8, pch =20,size =5, position=position_jitter(width=0.1))+
    theme_bw()+
    labs(x = 'Field Type', y=expression(paste("Average Res.  ",italic("Pythium")))) +
    theme(legend.position = "none")+
    theme(axis.title.x= element_blank(),
        axis.title.y = element_text(size =10, color ='black'),
        axis.text.x = element_text(size =10, color ='black'),
        axis.text.y = element_text(size =8, color = 'black'))+
scale_color_manual(name = 'Field Type', labels = c('Non-virgin', 'Virgin'), values = c('orange', 'chartreuse4'))+
  facet_wrap(~ Location)
#sb2
#dev.off()
```

### Fusarium

```{r}
options(repr.plot.width = 5, repr.plot.height = 3)
#jpeg('TuberwtAll_soilprints.jpg',width=5,height=3,units='in',res=300)
sb3= ggplot(data = df_21, aes(x =FieldType, y = Fus_Avg_Soil, color = FieldType))+
    geom_boxplot(width = 0.4)+
    geom_jitter(alpha=0.8, pch =20,size =5, position=position_jitter(width=0.1))+
    theme_bw()+
    labs(x = 'Field Type', y=expression(paste("Average ",italic("Fusarium")))) +
    theme(legend.position = "none")+
    theme(axis.title.x= element_blank(),
        axis.title.y = element_text(size =10, color ='black'),
        axis.text.x = element_text(size =10, color ='black'),
        axis.text.y = element_text(size =8, color = 'black'))+
scale_color_manual(name = 'Field Type', labels = c('Non-virgin', 'Virgin'), values = c('orange', 'chartreuse4'))+
  facet_wrap(~ Location)
#sb3
#dev.off()
```

### Verticillium

```{r}
options(repr.plot.width = 5, repr.plot.height = 3)
#jpeg('TuberwtAll_soilprints.jpg',width=5,height=3,units='in',res=300)
sb4= ggplot(data = df_21, aes(x =FieldType, y = Vd_Tot_Soil, color = FieldType))+
    geom_boxplot(width = 0.4)+
    geom_jitter(alpha=0.8, pch =20,size =5, position=position_jitter(width=0.1))+
    theme_bw()+
    labs(x = 'Field Type', y=expression(paste("Total ",italic("V. dahliae")))) +
    theme(legend.position = "none")+
    theme(axis.title.x= element_blank(),
        axis.title.y = element_text(size =10, color ='black'),
        axis.text.x = element_text(size =10, color ='black'),
        axis.text.y = element_text(size =8, color = 'black'))+
scale_color_manual(name = 'Field Type', labels = c('Non-virgin', 'Virgin'), values = c('orange', 'chartreuse4'))+
  facet_wrap(~ Location)
#sb4
#dev.off()
```

### Colletotrichum

```{r}
options(repr.plot.width = 5, repr.plot.height = 3)
#jpeg('TuberwtAll_soilprints.jpg',width=5,height=3,units='in',res=300)
sb5= ggplot(data = df_21, aes(x =FieldType, y = BD_Tot_Soil, color = FieldType))+
    geom_boxplot(width = 0.4)+
    geom_jitter(alpha=0.8, pch =20,size =5, position=position_jitter(width=0.1))+
    theme_bw()+
    labs(x = 'Field Type', y=expression(paste("Total ",italic("C. coccodes")))) +
    theme(legend.position = "none")+
    theme(axis.title.x= element_blank(),
        axis.title.y = element_text(size =10, color ='black'),
        axis.text.x = element_text(size =10, color ='black'),
        axis.text.y = element_text(size =8, color = 'black'))+
scale_color_manual(name = 'Field Type', labels = c('Non-virgin', 'Virgin'), values = c('orange', 'chartreuse4'))+
  facet_wrap(~ Location)
#sb5
#dev.off()
```

```{r}
options(repr.plot.width = 10, repr.plot.height = 6)
#jpeg('soil_pathogens_21.jpg',width=10,height=7,units='in',res=300)
plot = ggarrange(sb1,sb2,sb3,sb4,sb5, labels = c("A", "B", "C", "D","E"), ncol = 2, nrow = 3)
plot
#dev.off()
```


## Soil phyiscal and chemical properties

### Active Carbon

```{r}
options(repr.plot.width = 5, repr.plot.height = 3)
#jpeg('ActiveCarbon_21.jpg',width=5,height=3,units='in',res=300)
sp1= ggplot(data = df_21, aes(x =FieldType, y = POXC, color = FieldType))+
    geom_boxplot(width = 0.4)+
    geom_jitter(alpha=0.8, pch =20,size =5, position=position_jitter(width=0.1))+
    theme_bw()+
    labs(x = 'Field Type', y=expression(paste("Active Carbon (mg/kg) "))) +
    theme(legend.position = "none")+
    theme(axis.title.x= element_blank(),
        axis.title.y = element_text(size =10, color ='black'),
        axis.text.x = element_text(size =10, color ='black'),
        axis.text.y = element_text(size =8, color = 'black'))+
scale_color_manual(name = 'Field Type', labels = c('Non-virgin', 'Virgin'), values = c('orange', 'chartreuse4'))+
  facet_wrap(~ Location)
sp1
#dev.off()
```


# 2022 data visualization and inference

## Microplot Study

```{r}
yield = read_csv(file = 'C:/Users/theox/Desktop/Projects/SoilFingerPrints/Data/2022/Microplot/microplot_22.csv')
yield = yield[c('ID', 'location', 'pair', 'FieldType', 'TW_gt20', 'TW_All', 'TC_gt20', 'TC_All')]

yield$location = gsub('columbia Basin', 'Columbia Basin', yield$location)

yield$FieldType = gsub('Virgin field', "Virgin", yield$FieldType)
yield$FieldType = gsub('Non-virginfield', "Non-virgin", yield$FieldType)
yield$FieldType = gsub('Non-virgin Field', "Non-virgin", yield$FieldType)
yield$FieldType = gsub('Non-virgin field', "Non-virgin", yield$FieldType)
yield$FieldType = gsub('Virgin Field', "Virgin", yield$FieldType)

yield[c("ID", "pair", "FieldType","location")]  = lapply(yield[c("ID", "pair", "FieldType","location")], factor)
head(yield)
#unique(yield$FieldType)
#str(yield)
#lapply(yield, unique)
```
### Tuber weigt more than 20gm

```{r}
options(repr.plot.width = 5, repr.plot.height = 3)
#jpeg('Tuberwt_gt20_222_combinedlocations.jpg',width=5,height=3,units='in',res=300)
a= ggplot(data = yield, aes(x =FieldType, y = TW_gt20, color = FieldType))+
    geom_boxplot(width = 0.4)+
    geom_jitter(alpha=0.8, pch =20,size =5, position=position_jitter(width=0.1))+
    theme_bw()+
    labs(x = 'Field Type', y="Tuber weight (gm)") +
    theme(legend.position = "none")+
    theme(axis.title.x= element_blank(),
        axis.title.y = element_text(size =10, color ='black'),
        axis.text.x = element_text(size =10, color ='black'),
        axis.text.y = element_text(size =8, color = 'black'))+facet_wrap(~ location)+
scale_color_manual(name = 'Field Type', labels = c('Non-virgin', 'Virgin', 'Native'), 
                  values = c('tomato', 'orange', 'chartreuse4'))
#a
#dev.off()
```



### Tuber count more than 20gm

```{r}
options(repr.plot.width = 5, repr.plot.height = 20)
#jpeg('TuberCount_gt20_22_locationwise.jpg',width=5,height=3,units='in',res=300)
b= ggplot(data = yield, aes(x =FieldType, y = TW_gt20, color = FieldType))+
    geom_boxplot(width = 0.4)+
    geom_jitter(alpha=0.8, pch =20,size =5, position=position_jitter(width=0.1))+
    theme_bw()+
    labs(x = 'Field Type', y="Number of tubers") +
    theme(legend.position = "none")+
    theme(axis.title.x= element_blank(),
        axis.title.y = element_text(size =10, color ='black'),
        axis.text.x = element_text(size =10, color ='black'),
        axis.text.y = element_text(size =8, color = 'black'))+
scale_color_manual(name = 'Field Type', labels = c('Non-virgin', 'Virgin', 'Native'), 
                  values = c('tomato', 'orange', 'chartreuse4'))+facet_grid(~ location)
#b
#dev.off()
```

```{r}
options(repr.plot.width = 7, repr.plot.height = 6)
#jpeg('yield 2022_combinedLocations.jpg',width=7,height=6,units='in',res=300)
combined_plot = ggarrange(a,b, labels = c("A", "B"), ncol = 1, nrow = 2)
combined_plot
#dev.off()
```



### Tuber weigt more than 20gm

```{r, inclde = F, echo = F}
options(repr.plot.width = 5, repr.plot.height = 3)
#jpeg('Tuberwt_gt20_22_pairwise.jpg',width=5,height=3,units='in',res=300)
c= ggplot(data = yield, aes(x =pair, y = TW_gt20, color = FieldType))+
    geom_point(size = 2)+
    theme_bw()+
    labs(x = 'Field Type', y="Tuber weight (gm)") +
    #theme(legend.position = "none")+
    theme(axis.title.x= element_blank(),
        axis.title.y = element_text(size =10, color ='black'),
        axis.text.x = element_text(size =10, color ='black'),
        axis.text.y = element_text(size =8, color = 'black'))+
scale_color_manual(name = 'Field Type', labels = c('Native', 'Non-virgin', 'Virgin'), values = c('tomato', 'orange', 'chartreuse4'))+facet_wrap(~ location, scales = 'free')
#c
#dev.off()
```


### Tuber weigt more than 20gm

```{r, inclde = F, echo = F}
options(repr.plot.width = 5, repr.plot.height = 3)
#jpeg('Tubercount_gt20_22_pairwise.jpg',width=5,height=3,units='in',res=300)
d= ggplot(data = yield, aes(x =pair, y = TC_gt20, color = FieldType))+
    geom_point(size = 2)+
    theme_bw()+
    labs(x = 'Field Type', y="Tuber weight (gm)") +
    #theme(legend.position = "none")+
    theme(axis.title.x= element_blank(),
        axis.title.y = element_text(size =10, color ='black'),
        axis.text.x = element_text(size =10, color ='black'),
        axis.text.y = element_text(size =8, color = 'black'))+
scale_color_manual(name = 'Field Type', labels = c('Native', 'Non-virgin', 'Virgin'), values = c('tomato', 'orange', 'chartreuse4'))+facet_wrap(~ location, scales = 'free')
#d
#dev.off()
```


```{r, inclde = F, echo = F}
options(repr.plot.width = 7, repr.plot.height = 6)
#jpeg('yield 2022_pair and location wise.jpg',width=7,height=6,units='in',res=300)
combined_plot_pairwise = ggarrange(c,d, labels = c("A", "B"), ncol = 1, nrow = 2)
#combined_plot_pairwise
#dev.off()
```

### Inference (linear mixed model)
Here, locations: columbia basin and skagit valley are considered random effect and field type is considered fixed effect.

**Inference on tuber weight**

```{r}


fit_yield = lmer(TC_gt20 ~ FieldType + (1|location), 
                  data = yield)
#summary(fit_yield)
anova(fit_yield)

library(emmeans)
emmeans(fit_yield, list(pairwise ~ FieldType), adjust = "tukey")

library(lmerTest)
#difflsmeans(fit_yield, test.effs = "FieldType", df="Kenward-Roger")

#Multiple comparison
library(multcomp)
#summary(glht(fit_yield, linfct = mcp(FieldType = "Tukey")), test = adjusted("holm"))
```

```{r}
## for columbia basin
df_22_CB =yield[yield$location == 'Columbia Basin', ]
anova(lmer(TW_gt20 ~ FieldType + (1|pair), 
                  data = df_22_CB))
anova(lm(TW_gt20 ~ FieldType, data = df_22_CB))

#for skagit valley
df_22_SV = yield[yield$location == 'Skagit Valley', ]
fit_yield_22_SV = lmer(TW_gt20 ~ FieldType + (1|pair), 
                  data = df_22_SV)

#summary(fit_yield)
anova(fit_yield_22_SV)

library(lmerTest)
#difflsmeans(fit_yield_22_SV, test.effs = "FieldType", df="Kenward-Roger")

#Multiple comparison
library(multcomp)
summary(glht(fit_yield_22_SV, linfct = mcp(FieldType = "Tukey")), test = adjusted("holm"))

```



**Inference on tuber count**

```{r}
fit_count = lmer(TC_gt20 ~ FieldType + (1|location), 
                  data = yield)
#summary(fit_count)
anova(fit_count)
#plot(fit_count)

library(emmeans)
emmeans(fit_count, list(pairwise ~ FieldType), adjust = "tukey")
```

# 2022 Nematode Abundance 

## Data preparation

```{r}
df_NC = read_csv('C:/Users/theox/Desktop/Projects/SoilFingerPrints/Data/2022/Nematode_morpho/Nematode_raw_22.csv')

df_NC[c("Sites", "pair", "location", "FieldType")] <- lapply(df_NC[c("Sites", "pair", "location", "FieldType")],
                                                factor)

df_NC$FieldType = gsub('Field', "", df_NC$FieldType)
df_NC$FieldType = gsub('field', "", df_NC$FieldType)
df_NC$FieldType = gsub('Non-virgin ', "Non-virgin", df_NC$FieldType)
df_NC$FieldType = gsub('Virgin ', "Virgin", df_NC$FieldType)
df_NC$location = gsub('columbia Basin', 'Columbia Basin', df_NC$location)
unique(df_NC$FieldType)
unique(df_NC$location)

df_taxa = df_NC[-c(1,2,3,4)]
labels = df_NC[c(1,2,3,4)]
dim(df_NC)
```

## Visualization

```{r}

# check total abundance in each sample
rowSums(df_taxa)

# Turn percent cover to relative abundance by dividing each value by sample
# total abundance
comm <- decostand(df_taxa, method = "total")
# check total abundance in each sample
colSums(comm)
rowSums(comm)

# select major taxa using threshold 0.3

taxa_sel = comm[colSums(comm)>0.3]
names(taxa_sel)
taxa_other = comm[!colSums(comm)>0.3]
names(taxa_other)

names(taxa_other) = rep('Other',length(taxa_other))
dim(taxa_other); dim(taxa_sel)
#taxa_other = comm[!unlist(names(comm)) %in% sel_taxa]
```
* Melt to longer format

```{r}
df_melt = cbind(labels, taxa_sel, taxa_other)

df_melt = melt(as.data.table(df_melt), c("Sites", "location", "pair", "FieldType"))
dim(df_melt)
head(df_melt)
unique(df_melt$variable)
```
### plot Nematode Abundance all samples

```{r}
options(repr.plot.width = 15, repr.plot.height = 6)
#jpeg('Abundance_samples.jpg',width=12,height=6,units='in',res=300)
bar = ggplot(df_melt, aes(fill = variable, y=value, x= Sites)) + 
    geom_bar(position="fill", stat="identity", alpha = 1)+
  labs(x = 'Samples', y="Relative Abundance", title="Nematode Community 2022")+
    facet_wrap(~location, scales = 'free', ncol = 1)+theme(legend.title=element_blank())
bar
#dev.off()
```

### Nematode abundance based on field type

```{r}
options(repr.plot.width = 10, repr.plot.height = 6)
#jpeg('Abundance_fieldType.jpg',width=10,height=6,units='in',res=300)
bar = ggplot(df_melt, aes(fill = variable, y=value, x= FieldType)) + 
    geom_bar(position="fill", stat="identity", alpha = 1)+
  labs(x="Field type", y="Relative Abundance", title="Nematode Community 2022")+
  facet_wrap(~location, ncol = 1)+theme(legend.title=element_blank()) 
bar
#dev.off()
```

### Box plot

```{r}
options(repr.plot.width = 13, repr.plot.height = 7)
#jpeg('Boxplot_nematode abundance_fieldType.jpg',width=13,height=7,units='in',res=300)
boxplot = ggplot(df_melt, aes(y=value, x= variable, color = FieldType)) + 
    geom_boxplot(width = 0.9)+
    geom_jitter(alpha=0.8, pch =20,size =2, position=position_jitter(width=0.05))+
  labs(x="Major taxa", y="Relative Abundance", title="Nematode Community 2022")+ theme_bw()+
  theme(axis.text.x=element_text(angle =80, vjust=0.5, hjust=0.35), legend.title=element_blank())+
  facet_wrap(~location, ncol = 1, scales = 'free')+
  scale_color_manual(name = 'Field Type', labels = c('Native', 'Non-Virgin', 'Virgin'), 
                 values = c('tomato', 'orange', 'chartreuse4'))
  
boxplot
#dev.off()
```

## Shannon's diversity

```{r}

library(phyloseq)
library('picante')
```

```{r}
#Subsample reads
count_tab_phy <- otu_table(df_taxa, taxa_are_rows=F)
ps_rare <- rarefy_even_depth(count_tab_phy, rngseed = 123, replace = FALSE)
dim(ps_rare)
log_rare= log2(ps_rare +1)
shannon = phyloseq::estimate_richness(log_rare, measures = "Shannon")
#observed = phyloseq::estimate_richness(count_tab_phy, measures = "Observed")
diversity = cbind(shannon, labels)
head(diversity)
```

```{r}
options(repr.plot.width = 5, repr.plot.height = 20)
#jpeg('Alpha diveristy.jpg',width=5,height=3,units='in',res=300)
b= ggplot(data = diversity, aes(x =FieldType, y = Shannon, color = FieldType))+
    geom_boxplot(outlier.color = NA)+
    geom_jitter(aes(color = FieldType), height = 0, width = .2)+
    labs(x = 'Field Type', y="Shannon's diversity index") +
    theme(legend.position = "none")+
    theme(axis.title.x= element_blank(),
        axis.title.y = element_text(size =10, color ='black'),
        axis.text.x = element_text(size =10, color ='black'),
        axis.text.y = element_text(size =8, color = 'black'))+
scale_color_manual(name = 'Field Type', labels = c('Native', 'Non-Virgin', 'Virgin'), 
                  values = c('tomato', 'orange', 'chartreuse4'))+facet_wrap(~ location)
b
#dev.off()
```

**Inference**

```{r}
fit_shannon = lmer(Shannon ~ FieldType + (1|location), data = diversity)
summary(fit_shannon)

wilcox.observed = pairwise.wilcox.test(diversity$Shannon,
                     diversity$FieldType,
                     p.adjust.method = "BH")
wilcox.observed$p.value
```

## Beta diversity

### Ordination


```{r}
ord = metaMDS(log_rare, distance = 'bray')

options(repr.plot.width = 5, repr.plot.height = 3)
#jpeg('Beta_biplot.jpg',width=5,height=3,units='in',res=300)
beta = ggplot(labels, aes(x = ord$points[,1], y = ord$points[,2], 
                          color = FieldType, shape = location))+
  geom_point(size = 2)+
  labs(x = 'NMDS 1', y="NMDS 2") +
  theme(axis.title.x= element_text(size =10, color ='black'),
        axis.title.y = element_text(size =10, color ='black'),
        axis.text.x = element_text(size =10, color ='black'),
        axis.text.y = element_text(size =8, color = 'black'))+
  scale_color_manual(name = 'Field Type', labels = c('Native', 'Non-virgin', 'Virgin'), 
                  values = c('tomato', 'orange', 'chartreuse4'))+theme_bw()+
  stat_ellipse(aes(group = FieldType), linetype = 2)

beta
#dev.off()
```






```{r, include = F, echo = F}
ord = metaMDS(log_rare, distance = 'bray')

options(repr.plot.width = 5, repr.plot.height = 3)
#jpeg('Beta2.jpg',width=5,height=3,units='in',res=300)
beta = ggplot(labels, aes(x = ord$points[,1], y = ord$points[,2], color = FieldType))+
  geom_point(size = 2)+
  labs(x = 'NMDS 1', y="NMDS 2") +
  theme(axis.title.x= element_blank(),
        axis.title.y = element_text(size =10, color ='black'),
        axis.text.x = element_text(size =10, color ='black'),
        axis.text.y = element_text(size =8, color = 'black'))+
  scale_color_manual(name = 'Field Type', labels = c('Native', 'Non-virgin', 'Virgin'), 
                  values = c('tomato', 'orange', 'chartreuse4'))+
  facet_wrap(~location)+stat_ellipse(aes(group = FieldType), linetype = 2)
beta
#dev.off()
```



```{r, include = F, echo = F}
ord = metaMDS(log_rare, distance = 'bray')

options(repr.plot.width = 5, repr.plot.height = 3)
#jpeg('Tuberwt_gt20_222_combinedlocations.jpg',width=5,height=3,units='in',res=300)
beta = ggplot(labels, aes(x = ord$points[,1], y = ord$points[,2], color = FieldType, label= pair))+
  geom_point(size = 0.1)+
  geom_text(overlap = F)+
  labs(x = 'NMDS 1', y="NMDS 2") +
  theme(axis.title.x= element_blank(),
        axis.title.y = element_text(size =10, color ='black'),
        axis.text.x = element_text(size =10, color ='black'),
        axis.text.y = element_text(size =8, color = 'black'))+
  theme_bw()+
  scale_color_manual(name = 'Field Type', labels = c('Native', 'Non-virgin', 'Virgin'), 
                  values = c('tomato', 'orange', 'chartreuse4'))+
  facet_wrap(~location)
beta
#dev.off()
```



### Inference


```{r}
dist = phyloseq::distance(log_rare, method="bray")
permanova_pairwise<-adonis(dist ~ labels$FieldType, strata = labels$location)
permanova_pairwise$aov.tab
```


```{r}
#Generate distance matrix
clr_dist_matrix <- phyloseq::distance(log_rare, method = "euclidean") 

#ADONIS test
adon.results<-adonis(clr_dist_matrix ~ labels$FieldType, strata = labels$location)
adon.results$aov.tab

```



```{r}
#Dispersion test and plot
dispr <- vegan::betadisper(clr_dist_matrix, labels$FieldType)
plot(dispr, main = "Ordination Centroids and Dispersion Labeled: Aitchison Distance", sub = "")
```




```{r}
# and calculating our Euclidean distance matrix
euc_dist <- dist(log_rare)
length(euc_dist)
```

```{r}
euc_clust <- hclust(euc_dist, method="ward.D2")

#plot(euc_clust) # hclust objects like this can be plotted as is, but i like to change them to dendrograms for two reasons:
  # 1) it's easier to color the dendrogram plot by groups
  # 2) if you want you can rotate clusters with the rotate() function of the dendextend package

euc_dend <- as.dendrogram(euc_clust, hang=0.1)
plot(euc_dend, ylab="VST Euc. dist.")
```


















